Actuator line modeling of vertical-axis turbines

### 2016-05-13 paper.tex
Slim down end effects intro

Helmholtz's second vortex theorem states that vortex lines may not end in a
fluid, but must either form closed loops or extend to boundaries. Consequently
the lift distribution due to the bound vortex from foils of finite span must
drop to zero at the tips.

Glauert~\cite{Glauert1935} used Prandtl's lifting line theory \cite{Prandtl1927}
to develop a tip loss correction factor $F$ to be applied to the local lift of
an axial-flow rotor blade element. The Glauert correction was further refined
for horizontal-axis wind turbines by Shen et al. \cite{Shen2005a}. Despite their
success for the blade element analysis of axial-flow rotors, these corrections
both depend on rotor parameters---tip speed ratio, number of blades, element
radius, tip flow angle---that do no necessarily translate directly to the
geometry and flow environment of a cross-flow rotor. Therefore, a more general
end effects model was sought.

@@

We can therefore compute a correction function based on the normalized spanwise
lift coefficient distribution
\begin{equation}
    F = C_l(\theta)/C_l(\theta)_{\max},
\end{equation}
which will be in the range $[0, 1]$, similar to the Glauert corrections, but
does not contain rotor parameters.

>>>>

Helmholtz's second vortex theorem states that vortex lines may not end in a
fluid, but must either form closed loops or extend to boundaries. Consequently
the lift distribution due to the bound vortex from foils of finite span must
drop to zero at the tips. One popular end effects correction was developed by
Glauert~\cite{Glauert1935} for the blade element analysis of axial-flow rotors.
However this correction depends on rotor parameters---tip speed ratio, number of
blades, element radius, tip flow angle---that do no necessarily translate
directly to the geometry and flow environment of a cross-flow rotor. Therefore,
a more general end effects model was sought.

@@

We can therefore compute a correction function $F$ to be applied to the ALM lift
coefficient, based on the normalized spanwise lift coefficient distribution
\begin{equation}
    F = C_l(\theta)/C_l(\theta)_{\max},
\end{equation}
which will be in the range $[0, 1]$, similar to the Glauert corrections, but
does not contain rotor parameters.
